---
title: "Superior paper quality with R-Markdown: The OBVIOUS Study"
bibliography: refs.bib
csl: jama.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      dpi = 600, 
                      fig.width = 11, 
                      fig.height = 6)

# load standard packages for RMD word docs
library(officer)
library(flextable)
library(gtsummary)
library(dplyr)
library(purrr)

# specify a rounding specification 
library(table.glue)

conflicted::conflict_prefer('as_flextable', winner = 'flextable')
conflicted::conflict_prefer('filter', winner = 'dplyr')


rspec <- round_spec() |>
  round_using_magnitude(digits = c(2, 1, 0),
                        breaks = c(1, 10, Inf))

# save it to options:
names(rspec) <- paste('table.glue', names(rspec), sep = '.')
options(rspec)

# default is double space 
# single spacing for title page, abstract, captions
single_par <- fp_par(line_spacing=1)


```

Your name here, your degree,^1^ 
Second in Command, MS,^1^ 
Minor Contributor, PhD,^2^ and 
The Mastermind, PhD^1^ 

^1^Your department, University, City, State. `r single_par`
^2^Department of Win, Overtly Pompous University, Irvine, CA. `r single_par`

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

**Correspondence to:** \
Your name, Degree \
Your Department \
Additional departmental details if needed \
Your university \
Address line 1 \
Address line 2 \
Address line 3 \
your phone number \
your email `r single_par`

**Word Count:** Abstract: ; Manuscript: 

```{r data-prep, include=FALSE}

library(palmerpenguins)

data <- penguins

```

\newpage

# ABSTRACT

\newpage

# INTRODUCTION


Here's how to cite one thing.[@r_language]

Here's how to cite two or more things.[@ambrosius_design_2014; @sprint_research_group_randomized_2015]

# METHODS

# RESULTS

```{r results-prep}

n_penguins <- nrow(data)

mean_bill_length <- mean(data$bill_length_mm, na.rm = TRUE)

mean_mass <- mean(data$body_mass_g, na.rm = TRUE)
sd_mass <- sd(data$body_mass_g, na.rm = TRUE)

# make a short name for table.glue::table_value - this makes
# the inline reporting of results less complicated to read.
tv <- table.glue::table_value
tg <- table.glue::table_glue

```


Here's how to put some results into the text: A total of `r n_penguins` were included in the current analysis. The mean bill length was `r tv(mean_bill_length)`. The mean (standard deviation) body mass was `r tg("{mean_mass} ({sd_mass})")`

# DISCUSSION

\newpage

# REFERENCES

<div id="refs"></div>

\newpage

# ACKNOWLEDGMENTS 

# FINANCIAL DISCLOSURE 

```{r tbl_footers}

ckd_defn <- as_paragraph("Chronic Kidney Disease was defined by an estimated glomerular filtration rate <60 ml/min/1.73 m", as_sup(2L), " based on the 2021 CKD-EPI creatinine equation.")

```


```{r tbl_abbrvs}

abbrvs <- list(
  CI   = 'confidence interval',
  FI   = 'frailty index',
  SD   = 'standard deviation',
  IQR  = 'interquartile range',
  MoCA = 'Montreal cognitive assessment',
  EHR  = 'electronic health records',
  CVD  = 'cardiovascular disease'
)

abbrvs_paste <- function(strings){
  
  front <- 'Abbreviations'
  
  matter <- glue::glue_collapse(strings, sep = '; ', last = '; and ')
  
  paste(front, matter, sep = ': ')
  
}

abbrvs_write <- function(abbr){
  
  sorted <- abbr[sort(names(abbr))]
  
  strings <- map2_chr(names(sorted), sorted, paste, sep = ' = ')
  
  as_paragraph(abbrvs_paste(strings))
  
}


```

**Table 1: Penguin characteristics**

```{r}


tbl_summary(data, by = 'island') %>% 
 as_flex_table() %>% 
 theme_box() %>% 
 align(j = c(2,3,4), align = 'center')

```


\newpage

**Table 2: A more customized table**

```{r}

data %>% 
 group_by(species) %>% 
 slice(1:3) %>% 
 select(-body_mass_g, -sex, -year) %>% 
 mutate(across(where(is.numeric), table_value)) %>% 
 as_grouped_data(groups = 'species') %>% 
 as_flextable(hide_grouplabel = TRUE) %>% 
 set_header_labels(bill_length_mm = 'Bill length, mm',
                   bill_depth_mm = 'Bill depth, mm',
                   flipper_length_mm = 'Flipper length, mm') %>% 
 add_header_row(values = c(" ", " Penguin things"),
                colwidths = c(1, 3)) %>% 
 bg(i = ~!is.na(species), bg = 'grey90') %>% 
 width(width = 1.5) %>% 
 width(j = 1, width = 2) %>% 
 align(part = 'all', j = c(2,3,4), align = 'center') %>% 
 footnote(i = 2, j = 2, 
          part = 'header',
          value = as_paragraph("Bill length doesn't correlate with what you think it does."),
          ref_symbols = '1') %>% 
 footnote(i = 1, j = 1, 
          part = 'body',
          value = as_paragraph('Adelies are the best and they know it'), 
          ref_symbols = '2') %>% 
 add_footer_lines(values = abbrvs_write(abbrvs[c('CI', 'EHR', 'SD')])) %>% 
 add_footer_lines(values = "none of those abbreviations are in the table but they demonstrate how to include abbreviations using my helper functions")

```


<!-- new page not needed at the end of a block section -->

```{r}

margins_normal <- page_mar(bottom = 2.54 * 0.3937, 
                           top    = 2.54 * 0.3937, 
                           right  = 2.54 * 0.3937, 
                           left   = 2.54 * 0.3937,
                           gutter = 0)

block_section(
  prop_section(
    page_size = page_size(orient = "portrait"),
    type = "continuous",
    page_margins = margins_normal
  )
)

```

To make a landscape section, use the block_section code here. Note that the section will be applied to all content prior to the block reaching up to the most recent block.

```{r echo = TRUE}
block_section(
  prop_section(
    page_size = page_size(orient = "landscape"),
    page_margins = page_mar(bottom = 1/8, top = 1/8, 
                            right = 1/4, left = 1/4),
    type = "continuous"
  )
)
```


To switch back to regular portrait section, add another portrait block. Notice how the block below acts on all content above it up to the prior block.


```{r echo = TRUE}
block_section(
  prop_section(
    page_size = page_size(orient = "portrait"),
    type = "continuous",
    page_margins = margins_normal
  )
)
```

