---
title: "Individualized risk prediction for type 2 diabetes"
bibliography: refs.bib
csl: jama.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      dpi = 600, 
                      fig.width = 8, 
                      fig.height = 6,
                      fig.path = "doc/fig/")


# load standard packages for RMD word docs
library(officer)
library(flextable)
library(gtsummary)
library(dplyr)
library(purrr)
library(tidyverse)
library(survival)
library(broom)

# specify a rounding specification 
library(table.glue)

conflicted::conflict_prefer('as_flextable', winner = 'flextable')
conflicted::conflict_prefer('filter', winner = 'dplyr')


rspec <- round_spec() |>
  round_using_magnitude(digits = c(2, 1, 0),
                        breaks = c(1, 10, Inf))

# save it to options:
names(rspec) <- paste('table.glue', names(rspec), sep = '.')
options(rspec)

# default is double space 
# single spacing for title page, abstract, captions
single_par <- fp_par(line_spacing=1)

withr::with_dir(
  here::here(),
  code = {
    targets::tar_load(c(labels,
                        reviewer_1_cstats,
                        data_analysis,
                        analysis))
  }
)

margins_normal <- page_mar(bottom = 2.54 * 0.1,  #0.3937, 
                           top    = 2.54 * 0.1,  #0.3937, 
                           right  = 2.54 * 0.1,  #0.3937, 
                           left   = 2.54 * 0.1, #0.3937,
                           gutter = 0)

page_long_above <- function(){
  
  block_section(
    prop_section(
      page_size = page_size(orient = "portrait"),
      type = "continuous",
      page_margins = margins_normal
    )
  )
  
}

page_wide_above <- function(){
  block_section(
    prop_section(
      page_size = page_size(orient = "landscape"),
      page_margins = margins_normal,
      type = "continuous"
    )
  )
}

```

```{r tbl_abbrvs}

abbrvs <- list(
  CI   = 'confidence interval',
  FI   = 'frailty index',
  SD   = 'standard deviation',
  IQR  = 'interquartile range',
  CVD  = 'cardiovascular disease',
  DPP  = "Diabetes Prevention Program",
  MESA = "Multi-Ethnic Study of Atherosclerosis",
  NRI  = "Net reclassification index",
  IDI  = "Integrated discrimination index",
  AUC  = "Area underneath the receiver-operator characteristic curve",
  IPA  = "Index of prediction accuracy",
  HOMA = "Homeostatic Model Assessment"
)

abbrvs_paste <- function(strings){
  
  front <- 'Abbreviations'
  
  matter <- glue::glue_collapse(strings, sep = '; ', last = '; and ')
  
  paste(front, matter, sep = ': ')
  
}

abbrvs_write <- function(abbr){
  
  sorted <- abbr[sort(names(abbr))]
  
  strings <- map2_chr(names(sorted), sorted, paste, sep = ' = ')
  
  as_paragraph(abbrvs_paste(strings))
  
}

fn_mesa_baseline <- as_paragraph("Baseline for Multi-Ethnic Study of Atherosclerosis participants was visit 1 or 2 dependent on pre-diabetes status")

ft_tidy <- function(x){
  theme_box(x) %>% 
    align(align = 'center', part = 'all') %>%
    align(j = 1, align = 'left', part = 'all')
}


```

C-statistics (95% confidence interval) were `r reviewer_1_cstats$data_internal$glucose_fasting_mgdl` and 
`r reviewer_1_cstats$data_internal$hba1c_percent` for FPG and Hba1c, respectively, for internal validation. In external validation, the respective C-statistics were `r reviewer_1_cstats$data_external$glucose_fasting_mgdl` and 
`r reviewer_1_cstats$data_external$hba1c_percent`. Noting that the combination of FPG and Hba1c consistently obtained higher C-statistic and index of prediction accuracy compared to the individual variables, we replicated our main analysis with an additional model that was restricted to these two predictors (Table R1)



`r page_long_above()`

Table R1. Comparison of the individualized model with standard model.


```{r}

analysis$eval %>% 
  drop_na(reviewer_1) %>% 
  pivot_wider(names_from = evaluation, 
              values_from = c(standard, reviewer_1, individualized)) %>% 
  select(stat, 
         subgroup,
         reviewer_1_internal,
         standard_internal,
         individualized_internal,
         reviewer_1_external,
         standard_external,
         individualized_external) %>% 
  as_grouped_data('subgroup') %>% 
  as_flextable(hide_grouplabel = TRUE) %>%
  set_header_labels(
    stat = "Evaluation statistic",
    standard_internal = "Standard",
    reviewer_1_internal = "FPG + Hba1c",
    individualized_internal = "Individualized",
    standard_external = "Standard",
    reviewer_1_external = "FPG + Hba1c",
    individualized_external = "Individualized"
  ) %>% 
  add_header_row(values = c("Evaluation statistic", 
                            "Internal Validation", 
                            "External Validation"), 
                 colwidths = c(1, 3, 3)) %>% 
  ft_tidy() %>% 
  width(width = 1.5) %>% 
  merge_v(part = 'header', j = 1) %>% 
  add_footer_lines(
    values = abbrvs_write(abbr = abbrvs[c("AUC", "IPA")])
  ) %>% 
  footnote(
    i = 1,
    j = 1,
    part = 'header',
    value = as_paragraph('Table values are scaled by a factor of 100 for ease of interpretation.'),
    ref_symbols = '1'
  ) %>% 
  footnote(
    i = 1,
    j = 2,
    part = 'header',
    value = as_paragraph('Internal validation results are based on 10-fold cross-validation in the Diabetes Prevention Program data'),
    ref_symbols = '2'
  ) %>% 
  footnote(
    i = 1,
    j = 5,
    part = 'header',
    value = as_paragraph('External validation results are based on application of models fitted to the Diabetes Prevention Program data to the Multi-Ethnic Study of Atherosclerosis data.'),
    ref_symbols = '3'
  ) 
```

`r page_wide_above()`


```{r}

data_dpp <- data_analysis %>%
  filter(study == "DPP") %>% 
  mutate(age_cat = cut(age_yrs, breaks = c(0, 55, 65, Inf)))

fit_cph <- 
  coxph(Surv(time_diabetes, status_diabetes) ~ sex + education +
          chol_trig_mgdl + bmi + hba1c_percent + glucose_fasting_mgdl +
          treatment*age_yrs, data = data_dpp)

pvals_multivariable <- tidy(anova(fit_cph)) %>% 
  select(term, p.value) %>% 
  filter(str_detect(term, 'age_yrs')) %>% 
  mutate(term = c("main", "tx_interaction"),
         p.value = table_pvalue(p.value, drop_leading_zero = F)) %>% 
  deframe()

data_inc <- cmprsk::cuminc(ftime = data_dpp$time_diabetes,
                           fstatus = data_dpp$status_diabetes,
                           group = data_dpp$age_cat)

pval_unadjust <- data_inc$Tests %>% 
  as_tibble() %>% 
  mutate(pv = table_pvalue(pv, drop_leading_zero = FALSE))

data_gg <- data_inc
data_gg$Tests <- NULL

data_gg %<>%
  as.list() %>%
  map_dfr(~tibble(time = .x$time, est = .x$est), .id = 'age_group') %>%
  filter(time < 3) %>% 
  mutate(age_group = factor(age_group, 
                            levels = c("(0,55] 1",
                                       "(55,65] 1",
                                       "(65,Inf] 1"),
                            labels = c("< 55 years",
                                       "55 to 65 years",
                                       "> 65 years")))

```

We examined the cumulative incidence of diabetes by age groups (age < 55 years, age 55 to 65 years, and age > 65 years) in the DPP. We did not detect a difference in the cumulative incidence curves (Figure R1; p = `r pval_unadjust$pv`). We also did not detect an assocation between continuous age and incident diagnosis of diabetes in a Cox regression model with adjustment for sex, eduction, triglycerides, BMI, Hba1c, fasting plasma glucose, and randomized treatment (p = `r pvals_multivariable['main']`). 

\newpage

Figure R1: Cumulative incidence of diabetes diagnosis in the DPP by age group.

```{r}

ggplot(data_gg, aes(x=time, y=est, color = age_group)) +
  geom_line(linewidth = 1) + 
  theme_bw() + 
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Time, years", 
       y = "Cumulative incidence of diabetes, %",
       color = "Age group")

```

`r page_long_above()`
